Fast Regression of the Tritium Breeding Ratio in Fusion Reactors

### 2022-08-08 paper/results.tex
Removed textcolor and minor grammar adjustments.

The results displayed in~\Fref{fig:exp1-time-vs-reg} indicate that in the first,
simplified case GBTs clearly appear to be the most accurate as well as the
fastest surrogate family in terms of mean prediction time. Following that, we
note that ERTs, SVMs and ANNs also achieved satisfactory results with respect to
both examined metrics. In addition, prediction times of GBTs and SVMs show
relatively lower variance than those of ERTs and ANNs. Even though the remainder
of tested surrogate families does not exhibit prohibitive complexity, its
regression performance falls below average.

@@

Comparing these results with those of the second, unrestricted experiment (shown
in~\Fref{fig:exp2-time-vs-reg}), we observe that many surrogate families
consistently underperform. The least affected models appear to be GBTs, ANNs and
ERTs, which are known to be capable of capturing relationships involving mixed
feature types that were deliberately withheld in the first experiment. With only
negligible differences, the first two of these families appear to be tied for
the best performance as well as the shortest prediction time. We observe that
ERTs and RBFs also demonstrated satisfactory results, clearly outperforming the
remaining surrogates in terms of regression performance, and in some cases also
in prediction time.

@@

The results shown in~\Fref{fig:scaling} suggest that in terms of regression
performance the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as
more training points are introduced. While such families achieve nearly comparable
performance on the largest dataset, in the opposite case tree-based ensemble
approaches clearly outperform ANNs. For instance, GBTs achieve
$\text{MAE}=\num{0.107}$ that is nearly half of the $\text{MAE}=\num{0.186}$
reached by ANNs, representing a clear benefit given vastly disparate training
and prediction times. This trend continues to set sizes up to~\num{6000}.

@@

Consistent with our expectation, the shortest training times were achieved by
instance-based learning methods (KNN, IDW) that
are trained trivially at the expense of increased lookup complexity later during prediction.
Furthermore, we observe that the majority of tree-based ensemble algorithms also perform
and scale well, unlike RBFs and GPR which appear to behave superlinearly. We note that ANNs,
which are the only family to utilize parallelisation during training, show an
inverse scaling characteristic. We suspect that this effect may be caused
by a constant multi-threading overhead that dominates the training process
on relatively small sets.

Finally, all tested families with the exception of previously mentioned instance-based
models offer desirable prediction times. Analogous to previous experiments,
GBTs, ABTs and ANNs appear to be tied, as they not only exhibit
comparable times but also similar scaling slopes. Following those, we note a
clear hierarchy of ERTs, SVMs, GPR and RBFs, trailed by IDW and KNNs.

@@

	\caption{Regression performance of Models~1-8 (from left to right, top to
		bottom) in Experiment~4, viewed
		as true vs.~predicted TBR on a test set of a selected cross-validation
		fold. Points are coloured by density.}
	\label{fig:reg-performance}
\end{figure}

@@

Such tests revealed that while QASS has unmatched performance on its own
adaptively-sampled training set, it is outperformed by the baseline scheme on
uniformly-random evaluation sets. We suspected that while QASS excels in
learning the most strongly peaked regions of the TBR theory, this comes at the
expense of precision in broader, smoother regions where uniformly random
sampling suffices. Therefore a mixed scheme was implemented, with half MCMC
samples and half uniformly random samples incremented on each iteration, which
is also shown in~\Fref{fig:qasssampling}. An increase in initial sample size was also observed to improve precision in these smooth regions of the toy theory, as
the initial samples were obtained from a uniform random distribution. As shown
in ~\Fref{fig:qasssampling100k}, with~\num{100000} initial samples it was
possible to obtain a ${\sim}40\%$ decrease in error as compared to the baseline
scheme, from 0.0025 to 0.0015 mean averaged error. Comparing at the point of
termination for QASS, this corresponds to a ${\sim}6\%$ decrease in the number
of total samples needed to train a surrogate model while achieving the same error.

>>>>

The results displayed in~\Fref{fig:exp1-time-vs-reg} indicate that in the first,
simplified case GBTs clearly appear to be the most accurate as well as the
fastest surrogate family in terms of mean prediction time. Following that, we
note that ERTs, SVMs and ANNs also achieved satisfactory results with respect to
both examined metrics. In addition, prediction times of GBTs and SVMs show
relatively lower variance than those of ERTs and ANNs. Even though the remainder
of tested surrogate families do not exhibit prohibitive complexity, their
regression performance fall below the average.

@@

Comparing these results with those of the second, unrestricted experiment (shown
in~\Fref{fig:exp2-time-vs-reg}), we observe that many surrogate families
consistently underperform. The least affected models appear to be GBTs, ANNs and
ERTs, which are known to be capable of capturing relationships involving mixed
feature types that were deliberately withheld in the first experiment. With only
negligible differences, the first two of these families appear to be tied for
the best performance as well as the shortest prediction time. We observe that
ERTs and RBFs also demonstrated satisfactory results, clearly outperforming the
remaining surrogates in terms of regression performance, and in some cases also prediction time.

@@

The results shown in~\Fref{fig:scaling} suggest that in terms of regression
performance the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as
more training points are introduced. While such families achieve nearly comparable
performance on the largest dataset, in the opposite case tree-based ensemble
approaches clearly outperform ANNs. For instance, GBTs achieve
$\text{MAE}=\num{0.107}$, nearly half of the $\text{MAE}=\num{0.186}$
achieved by ANNs, representing a clear benefit given vastly disparate training
and prediction times. This trend continues for set sizes up to~\num{6000}.

@@

Consistent with our expectations, the shortest training times were achieved by
instance-based learning methods (KNN, IDW) that
are trained trivially at the expense of increased lookup complexity later during prediction.
Furthermore, we observe that the majority of tree-based ensemble algorithms also perform
and scale well, unlike RBFs and GPR which appear to behave superlinearly. We note that ANNs,
which are the only family to utilize parallelisation during training, show an
inverse scaling characteristic. We suspect that this effect may be caused
by a constant multi-threading overhead that dominates the training process
on relatively small sets.

Finally, all tested families with the exception of previously mentioned instance-based
models offered desirable prediction times. Analogous to previous experiments,
GBTs, ABTs and ANNs appeared to be tied, as they not only exhibited
comparable times but also similar scaling slopes. After those, we note a
clear hierarchy of ERTs, SVMs, GPR and RBFs, trailed by IDW and KNNs.

@@

	\caption{Regression performance of Models~1-8 (from left to right, top to
		bottom) in Experiment~4, viewed
		as true vs.~predicted TBR on a test set of a selected cross-validation fold. Points are colored by density.}
	\label{fig:reg-performance}
\end{figure}

@@

Such tests revealed that while QASS has unmatched performance on its own
adaptively-sampled training set, it is outperformed by the baseline scheme on
uniformly-random evaluation sets. We inferred that while QASS excels in
learning the most strongly peaked regions of the TBR theory, this comes at the
expense of precision in broader, smoother regions where uniformly random
sampling suffices. Therefore a mixed scheme was implemented, with half MCMC
samples and half uniformly-random samples incremented on each iteration, which
is also shown in~\Fref{fig:qasssampling}. An increase in initial sample size was also observed to improve precision in these smooth regions of the toy theory, as
the initial samples were uniformly-random. As shown
in~\Fref{fig:qasssampling100k}, with~\num{100000} initial samples it was
possible to obtain a ${\sim}40\%$ decrease in error as compared to the baseline
scheme, from 0.0025 to 0.0015 mean averaged error. Comparing at the point of
termination for QASS, this corresponds to a ${\sim}6\%$ decrease in the number
of total samples needed to train a surrogate model while achieving the same error.
